---
title: "UTAS report"
output: html_document
---
Enter a Comment
### Maneuver difficulty

Table 1. Percent failure per maneuver
```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
library(plyr)
library(repmis)
library(knitr)
library(zoo)
library(Hmisc)
ManeuverLevel<-repmis::source_data("www.dropbox.com/s/37ya7rfalh9b29q/SimulatorProc.csv",sep = "auto",header = TRUE)
perc.fail.maneuver<-ddply(ManeuverLevel,.(SubManeuver),summarise,perc.fail=100*sum(Grade==1)/length(SubManeuver))
perc.fail.maneuver<-perc.fail.maneuver[order(-perc.fail.maneuver$perc.fail),]
perc.fail.maneuver$perc.fail<-format(round(perc.fail.maneuver$perc.fail,2))
rownames(perc.fail.maneuver)<-NULL
kable(perc.fail.maneuver, format = "html")
```

### Analysis of Simulator Data per Maneuver

Objective: Find flight metrics that discriminate pilots' flight performance and 
associate to cockpit conditions and physiological responses.

###TCAS

In TCAS, changes in **calibrated airspeed** are strongly associated to instructors' grading (Ground truth?).

Absolute changes in air speed compared to air speed at the TCAS beginning: 

####Absolute delta calspeed=|CalibratedAirspeed (t) - CalibratedAirspeed (t=0)|
```{r echo=FALSE,message=FALSE,warning=FALSE}
SimulatorProc<-repmis::source_data("https://www.dropbox.com/s/3nbfwagh2bqw18m/ManeuverLevel.csv?dl=0",sep = ",",header = TRUE)
tcas<-subset(SimulatorProc,Maneuver=="TCAS")
tcas<-repmis::source_DropboxData("tcas.csv","ununbggz3uyvfsr",sep = ",",header = TRUE)
library(ggplot2)
ggplot(tcas,aes(x=systime2,y=delta.calspeed,color=factor(Grade)))+
geom_smooth(method="loess")+
theme_classic(base_family="Helvetica",base_size=12)+
  theme(legend.position="top")+labs(y="Abs Delta Airspeed [m/s?]",x="Time [ms]")+
  scale_color_manual(values=c("red", "green", "blue"),name="Instructor's\n Grade")
```

###Departure stall

Metric of interest: Speed recovery time


Description: time elapsed from minimum speed point to regaining 75% of the initial air speed 

Longer times were experienced by trials graded 1 and 2
```{r echo=FALSE,message=FALSE,warning=FALSE}
ggplot()+
  geom_point(data=subset(SimulatorProc,Maneuver=="Departure Stall"),alpha=0.1,aes(x=systime2,y=CalibratedAirspeed,group=Session))+
  facet_grid(Grade~.)+geom_vline(xintercept=1.65e5,color="red")+
  geom_vline(xintercept=1.3e5,color="green",alpha=0.5)+
  theme_bw(base_family="Helvetica",base_size=12)+
  annotate("text",x=1.1e5,y=260,label="start",size=4)+annotate("text",x=1.8e5,y=260,label="end",size=4,color="red")
```

###Upset Recovery

Metric: Altitude recovery after upset

Description: Approximately 50 seconds after the beginning of the maneuver, altitude values seem to stabilize near the initial altitude (~5000m?) . 

Deviations from this target altitude are the largest in the medium and high co2 conditions

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
ur<-subset(SimulatorProc,Maneuver=="Upset Recovery")
ur<-ur[order(ur$PID,ur$Session,ur$datetime),]
ur<-ddply(ur,.(PID,Session),transform, roll.sd.bankangle=rollapply(BankAngle,width=10,align="center",partial=TRUE,FUN=sd))



ggplot()+
  geom_line(data=subset(SimulatorProc,Maneuver=="Upset Recovery"),aes(x=systime2,y=Altitude,group=PID,color=factor(Grade)))+
  #geom_point(data=subset(manx2,Maneuver=="Upset Recovery"&Alarm1==1),aes(x=systime2,y=Sideslip+10),color="purple",size=1.2)+
  facet_grid(Condition~.)

kable(ddply(subset(SimulatorProc,Maneuver=="Upset Recovery"&systime2>50e3),.(Condition),summarise,Mean_altitude=mean(abs(Altitude),na.rm=TRUE)),format="html")


```

Metric: Bank Angle variability

Description: bank angle values oscillate closer to zero degrees 50 seconds after the beginning of the maneuver.

Absolute magnitude and deviation are larger in the medium and high CO2 conditions.





