---
title: "UTAS report"
output: html_document
---

### Maneuver difficulty

Table 1. Percent failure per maneuver
```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
library(plyr)
library(knitr)
library(zoo)
library(Hmisc)
library(readr)
setwd("..")
ManeuverLevel<-read_csv(gsub(" ","",paste(getwd(),"/Data/ManeuverLevel.csv")))
perc.fail.maneuver<-ddply(ManeuverLevel,.(SubManeuver),summarise,perc.fail=100*sum(Grade==1)/length(SubManeuver), Difficulty=Difficulty[1])
perc.fail.maneuver<-perc.fail.maneuver[order(-perc.fail.maneuver$perc.fail),]
perc.fail.maneuver$perc.fail<-format(round(perc.fail.maneuver$perc.fail,2))
rownames(perc.fail.maneuver)<-NULL
kable(perc.fail.maneuver, format = "html")
```

### Analysis of Simulator Data per Maneuver

Objective: Find flight metrics that discriminate pilots' flight performance and 
associate to cockpit conditions and physiological responses.

###TCAS

In TCAS, changes in **calibrated airspeed** are strongly associated to instructors' grading (Ground truth?).

Absolute changes in air speed compared to air speed at the TCAS beginning: 

####Absolute delta calspeed=|CalibratedAirspeed (t) - CalibratedAirspeed (t=0)|
```{r echo=FALSE,message=FALSE,warning=FALSE}
setwd("..")
SimulatorProc<-read_csv(gsub(" ","",paste(getwd(),"/Data/SimulatorProc2.csv")))
SimulatorProc$BankAngle <- as.numeric(SimulatorProc$BankAngle)
SimulatorProc$AltitudeDev <- as.numeric(SimulatorProc$AltitudeDev)
tcas<-subset(SimulatorProc,Maneuver=="TCAS")
#tcas<-repmis::source_DropboxData("tcas.csv","ununbggz3uyvfsr",sep = ",",header = TRUE)
library(ggplot2)
##MEMO - Fix delta.calspeed
#ggplot(tcas,aes(x=systime2,y=delta.calspeed,color=factor(Grade)))+
#geom_smooth(method="loess")+
#theme_classic(base_family="Helvetica",base_size=12)+
#  theme(legend.position="top")+labs(y="Abs Delta Airspeed [m/s?]",x="Time [ms]")+
#  scale_color_manual(values=c("red", "green", "blue"),name="Instructor's\n Grade")
```

###Departure stall

Metric of interest: Speed recovery time

Description: time elapsed from minimum speed point to regaining 75% of the initial air speed 

Relation to Instructor Ratings:

Longer times were experienced by trials graded 1 and 2
```{r echo=FALSE,message=FALSE,warning=FALSE}
ggplot()+
  geom_point(data=subset(SimulatorProc,Maneuver=="Departure Stall"),alpha=0.1,aes(x=systime2,y=CalibratedAirspeed,group=Session))+
  facet_grid(Grade~.)+geom_vline(xintercept=1.65e5,color="red")+
  geom_vline(xintercept=1.3e5,color="green",alpha=0.5)+
  theme_bw(base_family="Helvetica",base_size=12)+
  annotate("text",x=1.1e5,y=260,label="start",size=4)+annotate("text",x=1.8e5,y=260,label="end",size=4,color="red")
```

Relation to CO2 Conditions:

YY

###Upset Recovery

Metric: Altitude recovery after upset

Description: Approximately 50 seconds after the beginning of the maneuver, altitude values seem to stabilize near the initial altitude (~5000m?) . 

Deviations from this target altitude are the largest in the medium and high co2 conditions

Relation to Instructor Ratings:

YY

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
ur<-subset(SimulatorProc,Maneuver=="Upset Recovery")
ur<-ur[order(ur$PID,ur$Session,ur$datetime),]
ur<-ddply(ur,.(PID,Session),transform, roll.sd.bankangle=rollapply(BankAngle,width=10,align="center",partial=TRUE,FUN=sd))

ggplot()+
  geom_line(data=subset(SimulatorProc,Maneuver=="Upset Recovery"),aes(x=systime2,y=Altitude,group=PID,color=factor(Grade)))+
  #geom_point(data=subset(manx2,Maneuver=="Upset Recovery"&Alarm1==1),aes(x=systime2,y=Sideslip+10),color="purple",size=1.2)+
  facet_grid(Condition~.)

kable(ddply(subset(SimulatorProc,Maneuver=="Upset Recovery"&systime2>50e3),.(Condition),summarise,Mean_altitude=mean(abs(Altitude),na.rm=TRUE)),format="html")
```

Metric: Bank Angle variability

Description: bank angle values oscillate closer to zero degrees 50 seconds after the beginning of the maneuver.

Absolute magnitude and deviation are larger in the medium and high CO2 conditions.

###Circle to Land

Metric: Maintain altitude during circle to land

Description: Absolute deviation from 700 ft during circle to land

Relation to Instructor Ratings:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
kable(ddply(subset(SimulatorProc,SubManeuver=="Circle to Land" & is.na(Grade)==F),.(Grade),summarise,Mean_Deviation=mean(AltitudeDev,na.rm=TRUE)),format="html")
```

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
CircleToLand <- subset(SimulatorProc,SubManeuver=="Circle to Land")
ManeuverTime = function(x){
  x$systime3 = x$SysTime-x$SysTime[1]
  return(x)
}
CircleToLand <- ddply(CircleToLand,.(ManeuverNum), .fun=ManeuverTime)
ggplot()+ xlim(0,120000)+
  stat_smooth(data=CircleToLand, aes(x=systime3,y=AltitudeDev,color=factor(Condition)))
```

###Takeoff

Metric: Bank angle during takeoff

Description: Absolute deviation from 0 during takoff, up to 400 ft

Relation to Instructor Ratings:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
Takeoff <- subset(SimulatorProc,SubManeuver=="Takeoff")
Takeoff <- ddply(Takeoff,.(ManeuverNum), .fun=ManeuverTime)
ggplot()+ xlim(0,50000)+
  stat_smooth(data=subset(Takeoff, is.na(Grade)==F),aes(x=systime3,y=BankAngleDev,color=factor(Grade)))
```

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
ggplot()+ xlim(0,50000)+
  stat_smooth(data=Takeoff,aes(x=systime3,y=BankAngleDev,color=factor(Condition)))
```

###Area Departure with Engine Fire

Metric: Constant Rate of Ascent from 400ft to 1500ft

Description: Absolute deviation from pitch attitude of 15

Relation to Instructor Ratings:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
AreaDeparture <- subset(SimulatorProc,SubManeuver=="Area Departure")
AreaDeparture <- ddply(AreaDeparture,.(ManeuverNum), .fun=ManeuverTime)
ggplot()+ xlim(0,50000)+
  stat_smooth(data=subset(AreaDeparture, is.na(Grade)==F & Maneuver=="Engine Fire"),aes(x=systime3,y=Altitude,color=factor(Grade)))

kable(ddply(subset(AreaDeparture, is.na(Grade)==F & Maneuver=="Engine Fire"),.(Grade),summarise,Mean_Deviation=mean(PitchAttitudeDev,na.rm=TRUE)),format="html")
```

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
kable(ddply(subset(AreaDeparture, is.na(Grade)==F & Maneuver=="Engine Fire"),.(Condition),summarise,Mean_Deviation=mean(PitchAttitudeDev,na.rm=TRUE)),format="html")
```

###Steep Turns

Metric: Bank Angle of 45 Degrees

Description: Deviation from 45 degrees while in steep turns

Relation to Instructor Ratings:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
SteepTurns <- subset(SimulatorProc,SubManeuver=="Steep Turns")
SteepTurns <- ddply(SteepTurns,.(ManeuverNum), .fun=ManeuverTime)
kable(ddply(SteepTurns,.(Grade),summarise,Mean_Deviation=mean(BankAngleDev,na.rm=TRUE)),format="html")
```

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
ggplot()+ xlim(0,140000)+
  stat_smooth(data=SteepTurns,aes(x=systime3,y=BankAngle,color=factor(Condition)))

kable(ddply(SteepTurns,.(Condition),summarise,Mean_Deviation=mean(BankAngleDev,na.rm=TRUE)),format="html")
```

###Landing

Metric: Bank angle during final approach and landing

Description: Absolute deviation from 0 during landing

Relation to Instructor Ratings:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
Landing <- subset(SimulatorProc,SubManeuver=="Landing")
ManeuverTime = function(x){
  x$systime3 = x$SysTime-x$SysTime[1]
  return(x)
}
Landing <- ddply(Landing,.(ManeuverNum), .fun=ManeuverTime)
ggplot()+ xlim(0,50000)+
  stat_smooth(data=subset(Landing, is.na(Grade)==F),aes(x=systime3,y=BankAngleDev,color=factor(Grade)))
```

Relation to CO2 Conditions:

```{r echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
ggplot()+ xlim(0,50000)+
  stat_smooth(data=Landing,aes(x=systime3,y=BankAngleDev,color=factor(Condition)))
```

